
# Frequent Problems

## First exploration of new dataset

The `skim()`shows how many missing and unique values each variable has. It uses appropriate measures to describe each variable based on its type: character, numeric or list.

```{r}
skimr::skim(starwars)
```

The glimpse function, on the other hand, gives us a good peak at the first raw values each variable has.

```{r}
glimpse(starwars)
```

## A Count and prop table

**First way with `forcats::fct_count()`** Calculates a count and prop table.

```{r}
starwars$sex %>%
 factor() %>% 
  fct_count(sort = T, prop = T)
```

**Second way with `deplyr::count()`** Simply mutate a frequency and percentage column on a counted table.

```{r}
starwars %>% 
 count(sex) %>% 
  mutate(freq = n / sum(n)) %>% 
  mutate(perc = freq * 100)
```

## Bar graph with count data

Here is a situation where we calculated a count table for hair color - we summarized all values. If we then want to plot a bar graph based on that count table we run into problems, because ggplot2 is expecting a non-summarized or normal data frame.

```{r}
hair_color_table = starwars %>% 
  mutate(hair_color = fct_lump_min(hair_color, 2)) %>% 
  group_by(hair_color) %>% 
  summarise(n = n())
hair_color_table
```

To tell the function that we have already summarized data, we add the argument `stat = "identity"` to the `geom_bar()` function.

```{r}
hair_color_table %>% 
  ggplot(aes(x = reorder(hair_color, n), y = n, fill = hair_color)) + 
  geom_bar(stat = "identity") + 
  theme(legend.position = "none")
```

## Bar graph with percentage labels

First we create a table with counts and percentages:

```{r}
d = starwars %>% 
  group_by(gender) %>% 
  summarise(count = n()) %>% 
  mutate(percentage = count/sum(count))
d
```

Then we plot a graph with bar and with percentage labels.

```{r}
d %>% 
  ggplot(aes(gender, percentage, label = round(percentage, 2), fill = gender)) + 
  geom_bar(stat = "identity") + 
  geom_label(aes(fill = NA), fill = "white") + 
  theme(legend.position = "none")
```

## Collapse factors to „Other"

This syntax mutates the categorical variable homeworld into eight of its most frequent values. The other values are being collapsed into the categorical value „other".

```{r}
starwars %>% 
  mutate(homeworld = fct_lump_n(homeworld, n = 8)) %>% 
  group_by(homeworld) %>% 
  summarise(mean(height, na.rm =T), mean(mass, na.rm = T), n())
```

## Filter for specific values

We can easily filter out cases with certain column values, like for example the states of Hawai and Alaska. We use `filter()`, the operator `!` and `%in%`.

```{r}
starwars %>% 
  filter(!homeworld%in%c("Tatooine","Naboo")) %>% 
  select(name, homeworld)
```

## Change bar colors in barplot

You can manually pick the colors with `fill` and a vector containing the color values. Either in String, written out.

```{r}
starwars %>% 
  mutate(sex = fct_infreq(sex)) %>% 
  ggplot(aes(sex)) + 
  geom_bar(fill = c("red","blue","green","black","grey")) 
```

Or with RGB Color Codes.

```{r}
starwars %>% 
    mutate(sex = fct_infreq(sex)) %>%
    ggplot(aes(sex)) +
    geom_bar(fill = c("#003f5c","#58508d","#bc5090","#ff6361","#ffa600")) 
```

## Hide aes(color) mapping legend

Here is an example where we want the bar colored based on the variable itself, but without the mapping legend.

```{r}
starwars %>% 
  mutate(sex = fct_infreq(sex)) %>% 
  ggplot(aes(sex, fill = sex)) + 
  geom_bar()
```

Hide the geom_bar legend.

```{r}
starwars %>% 
  mutate(sex = fct_infreq(sex)) %>% 
  ggplot(aes(sex, fill = sex)) + 
  geom_bar(show.legend = F)
```

Remove just the legend title:

```{r}
starwars %>% 
  mutate(sex = fct_infreq(sex)) %>% 
  ggplot(aes(sex, fill = sex)) + 
  geom_bar() +
  theme(legend.title = element_blank())
```

Hide all legends created:

```{r}
starwars %>% 
  mutate(sex = fct_infreq(sex)) %>% 
  ggplot(aes(sex, fill = sex)) + 
  geom_bar() +
  theme(legend.position = "none")
```

## Re-code values of categorical variables

**First way** We can use `fct_collapse()`to create a new column with the new recoded values in it.

```{r eval=FALSE, include=FALSE}
ess_9 = ess_9 %>% 
  mutate(new_wltdffr = fct_collapse(wltdffr, 
             Too_Large = c("Large, extremely unfair", "Large, very unfair", "Large, somewhat unfair","Large, slightly unfair"), 
             Fair = "Fair", 
             Too_Small = c("Small, extremely unfair", "Small, very unfair", "Small, somewhat unfair","Small, slightly unfair"))) %>% 
  mutate(new_wltdffr = factor(new_wltdffr, levels = c("Too_Large", "Fair", "Too_Small")))
```

**Second way** By using `mutate`, to create a new column with our own values and `case_when`, to run through our observations looking for defined cases, together with "variable" `%in%`, we can create our own groups.

```{r}
gapminder %>% 
 mutate(group = case_when(
    region %in% c("Western Europe", "Northern Europe","Southern Europe","Northern America", "Australia and New Zealand") ~ "West", # If region is one of values -> assign it "West" in new group column.
    region %in% c("Eastern Asia", "South-Eastern Asia") ~ "East Asia",
    region %in% c("Caribbean", "Central America", "South America") ~ "Latin America",
    continent == "Africa" & 
      region != "Northern Africa" ~ "Sub-Saharan",
    TRUE ~ "Others")) %>%  # If nothing above applies -> assign it "Others" in group column
  head(10)

```

We turn this `group` variable into a factor to control the order of the levels:

```{r eval=FALSE, include=FALSE}
gapminder %>% 
  mutate(group = factor(group, levels = c("Others", "Latin America", "East Asia", "Sub-Saharan","West")))
```

## Order color legend

Order color legend by a variable's values.

```{r eval=FALSE, include=FALSE}
data %>% 
    ggplot(aes(year,population, color = reorder(country, desc(population)))) + # "Reorder countries by descending order of population"
    geom_line() 
```

## Show unique values

Display all unique values of variable.

```{r}
distinct(starwars, species) # dplyr function
```

**Note**: `distinct(dat$countries)` doesn't work.

## Slice rows by maximum or minimum values

**Note:** parameter `n` must be explicitly written, otherwise it throws an error.

```{r}
starwars %>% 
  slice_max(height, n = 5)
```

Show me 5% of the lowest height rows.

```{r}
starwars %>% 
  slice_min(height, prop = 0.05)
```

## Show Number of `NA`s

For a quick check of how many missing values there are in a single column:

```{r}
sum(is.na(starwars$height))
```

And how many are not `NA`s.

```{r}
sum(!is.na(starwars$height))
```

For a more detailed overview of the whole dataset use `skim()`. It shows a very useful `complete_rate` which tells us how much of the column is disturbed by missing values.

```{r}
skimr::skim(starwars)
```

## Drop rows with missing values

Drop rows that have `NA`values in a specific column, here in `height`.

```{r}
starwars %>% 
  drop_na(height)
```

Drop all rows that contain `NA` in any column.

```{r}
starwars %>% 
  drop_na()
```

Filter out any NA containing rows.

```{r}
starwars %>% 
 na.exclude()
```

## Replace `NA`s

Replace 0 with value you want as a replacement.

```{r}
data(na_example)
sum(is.na(na_example))

no_nas <- ifelse(is.na(na_example), 0, na_example) # "if is NA is true, change value to 0, else keep the value (i.e. na_example)"

sum(is.na(no_nas))
```

## The factor variable trap

The FVT is about what happens when you try to return factorized vectors into numeric values. Let's look at this with this code.

```{r}
z <-factor(c("12", "13", "14", "15", "12")) # We create an object by directly factorizing a vector. 
z

y <- as.numeric(z) # Now we want to convert them into numeric values. 
y # What?
```

This happened, because we picked up the on the factorization result. `factor()` assigns every element, based on its value, an integer number.

```{r}
typeof(z) # 1=12, 13=2, 14=3, 15=4, 12=1
```

To fix this problem, first convert the object back to character and then to numeric.

```{r}
y <- as.numeric(as.character(z))
y
```
