# R Playgrounds

```{r include=FALSE}
library(tidyverse)
library(dslabs)
```



## Expected Value and Standard Deviation

Roulette table example with Red, Black and Green. Only Green wins. Whats the probability that you get Green?

```{r}
v = rep(c("Red", "Black", "Green"), c(18,18,2))
prop.table(table(v))
```

This is the sampling model.

```{r}
x = sample(c(17,-1), prob = c(2/38,36/38))
```

The expected value is calculated by adding the possible value times their likelyhood together. Its formula is ap + b(1-p). The expected value for 1000 draws.

```{r}
EV = 1000 * (17*2/38 + (-1*36/38))
```

The standard error (standard deviation of random variables i.e., probability distributions). Its formula is a - b \* sqrt(p \* (1-p)). The standard error for 1000 draws.

```{r}
SE = sqrt(1000) * (-((-1)-17) * sqrt(2/38 * 36/38))
```

Random variable S storing the experimental values from sampling model.

```{r}
set.seed(1)
S = sample(c(17,-1), size = 1000, replace = T, prob = c(2/38, 36/38))

sum(S)
```

Create the experimental sampling distribution of the sample sum.

```{r}
roulette_winnings = function(){
  S = sample(c(17,-1), size = 1000, replace = T, prob = c(2/38, 36/38))
  sum(S)
}

set.seed(1)
S = replicate(10000, roulette_winnings())
hist(S)
```

The mean or expected value of X?

```{r}
mean(S)
```

The standard deviation or standard error of X?

```{r}
sd(S)
```

The probabilty that we win?

```{r}
mean(S > 0)
n = 1000
pbinom(500, size = 1000, prob = 1/19) # ??
```

## Theoretical and Experimental Standard Error

The sample statistics: 0.45 "democrats and a sample size of 100.

```{r}
p = 0.45

n = 100

```

The theoretical standard error for p = 0.45 and sample size n = 100.

```{r}
SE = sqrt(p * (1 - p)) / sqrt(n)
SE
```

Now we can experimentally proof this expected standard error by running a Monte Carlo simulation. We basically create a sampling distribution of size 10000 of the sample proportions with p = 0.45 and n = 100. Only that we lastly subtract p to find out the actual error.

```{r}
test_errors = replicate(10000, mean(sample(c(1,0), replace = T, size = n, prob = c(p, (1 - p)))))

# Distribituon of errors
test_errors = test_errors - p

# The standard deviation of the errors
sd(test_errors)

# Very close to the theoretical standard error
SE
```

And the error distribution is approximatly normal.

```{r}
qqnorm(test_errors);qqline(test_errors)
```

## Sample Size Resulting in 0.01 SE

The maximum SE in relation to proportions is with p = 0.5. Therefore we will take this worst case scenario case to calculate our goal, the required sample size to get a standard error arround 0.01. We can calculate a list of SE based on sample sizes 100 to 5000.

```{r}
p = 0.5
n = 100:5000

list_of_SEs = sqrt(p * (1 - p) / n)
head(list_of_SEs, 20)
```

When plotting the list of standard errors we can see at what sample size we will reach a standard error of around 0.01: with a sample size of around 2.500.

```{r}
plot(list_of_SEs, type = "l")
```

## Bank and Loans

```{r}
# Number of loans
n = 10000

# Probability of default
p = 0.03

# Loss per single forclosure
loss_per_forclosure = -200000

# Interest_rate
x = 0

# Random variable S storing defaults = 1 and non defaults = 0
S = sample(c(0,1), prob = c(1-p, p), size = n, replace = T)
head(S, 100)

# Expected value for 10000 loans 
sum(S * loss_per_forclosure)

# Monte-Carlo simulation amount 
B = 10000

# Simulation
losses = replicate(B, { 
  S = sample(c(0,1), prob = c(1-p, p), size = n, replace = T)
  sum(S * loss_per_forclosure)
})

# Losses distribution from simulation
hist(losses / 10^6)

# Expected value of S
EV = x*(1-p) + loss_per_forclosure*p
EV

# Standard error of S
SE = abs(x - loss_per_forclosure) * sqrt(p * (1-p))
SE

# Loans of 180000
x = 180000

# loss_per_forclosure*p + x * (1-p) = 0
# Find out x!
x = - (loss_per_forclosure * p / (1 - p))
x

# Interest of 6185.57 dollars for each loan needed to get on average 0 in total back as the bank.
EV = loss_per_forclosure * p + x * (1 - p) 
EV # Correct!

```

## Real Polling Data

```{r}
# library(dslabs)
data("polls_us_election_2016")

# Exclude observations that are too old.  
polls <- polls_us_election_2016 %>% 
  filter(enddate >= "2016-10-31" & state == "U.S.") 

head(polls, 5)
```

The first poll. Create a confidence interval.

```{r}
n = polls$samplesize[1]
x_hat = polls$rawpoll_clinton[1]/100



se_hat = sqrt(x_hat * (1 - x_hat) / n)

cf = c(x_hat - 1.96 * se_hat, x_hat + 1.96 * se_hat)
rm(x_hat, se_hat)
```

Create columns for x_hat, se_hat, lower and upper confidence bounds. Select only the relevant columns.

```{r}
polls = polls %>% 
  mutate(x_hat = polls$rawpoll_clinton/100, 
         se_hat = sqrt(x_hat * (1 - x_hat) / samplesize),
         lower = x_hat - 1.96 * se_hat, 
         upper = x_hat + 1.96 * se_hat) %>% 
  select(pollster, enddate, x_hat, se_hat, lower, upper)
```

Create a `hit` column indicating whether our confidence intervals included our true parameter, the final vote count for Clinton 48.2.

```{r}
polls = polls %>% 
  mutate(hit = ifelse(0.482 > lower & 0.482 < upper, TRUE, FALSE))

mean(polls$hit)
```
